Hurricane Evacuation Strategies for the State of Mississippi

### 2015-11-16 manuscript.tex
done executive summary

\section{Executive Summary}
\label{sec:summary}

\section{Introduction}
\label{sec:introduction}
  The Mississippi Emergency Management Administration (MSEMA) has hired our team to design hurricane evacuation routes for at-risk counties in the coastal regions of southern Mississippi. Given that this coastal area is a strong commercial center, the government authorities want us to minimize the number of unnecessary evacuations. Under this requirement, it is more advantageous to withhold ordering an evacuation for as long as safely possible. The reasons for this are two-fold. First, this ensures that local businesses can remain open for as long as possible. Second, this ensures that the forecasts at the time of the evacuation decision are as accurate as possible. Therefore, our models seek to recommend an optimal evacuation plan as well as the latest possible time at which it is safe to enact this plan.

>>>>

\section{Executive Summary}
\label{sec:summary}
Governor Bryant,\\

At the request of the Mississippi Emergency Management Administration, we have conducted a thorough mathematical analysis of hurricane evacuation strategies for all 82 counties in the state of Mississippi. These strategies were tested across a range of possible hurricane conditions backed by historical weather data. By testing across this range of possibilities, we found that it is best to follow one of two different strategies depending on the type of storm being faced.\\

In the event of a moderately sized, category 1, 2, or 3 hurricane, we recommend a strategy of evacuating every county in the coastal region. These evacuations should head in a direction that is most pointed away from the expected landfall. If many of the direction choices are very similar, these counties should prefer their north-west neighbors, as hurricanes from the gulf tend to curve toward the north-east. The virtues of this strategy are that it's relatively localized and allows each county to independently pick it's evacuation direction. As a result of this, evacuations of this type are very quick to organize, and allow for a rapid evacuation of the most endangered areas. This is useful for moderately sized hurricanes because it allows you to take more time in deciding whether or not to evacuate. You can wait to see if the storm will die down before hitting the coast, but are still able to spring into action if it doesn't.\\

In the event of a very large, category 4 or 5 hurricane, we recommend an orderly evacuation of the entire lower half of Mississippi toward the northern border. If a storm of this type appears on the radar, it is very unlikely to die down before hitting the coast, and is more likely to pass over Mississippi with its increased size. Strong governmental organization is required as early as possible to prevent a catastrophic loss of life. We have code capable of solving for the optimal allocation of traffic on each highway to maximize the throughput of the entire evacuation. This is a very heavy-handed response that should be reserved for the most dire of circumstances. In this event that this is needed, we encourage you to consult us for a solution tailored to that specific storm.\\

While these two strategies can aid you in making decisions for an imminent hurricane, we also have a recommendation for a long-term improvement to all evacuation routes. In our analysis of evacuation throughput, we discovered that the area around George County poses a significant bottleneck for highway traffic. As our analysis will show later in this manuscript, adding an additional highway route from George County to Perry county will significantly shift this bottleneck upward. This would allow more evacuees to escape the most dangerous coastal areas before they are slowed down by a bottleneck. Especially with potentially heavy traffic coming up highway 59 from New Orleans, traffic congestion in the coastal region is a problem requiring a more long term solution.

\section{Introduction}
\label{sec:introduction}
  The Mississippi Emergency Management Administration (MSEMA) has hired our team to design hurricane evacuation routes for at-risk counties in the coastal regions of southern Mississippi. Given that this coastal area is a strong commercial center, the government authorities want us to minimize the number of unnecessary evacuations. Under this requirement, it is more advantageous to withhold ordering an evacuation for as long as is safely possible. The reasons for this are two-fold. First, this ensures that local businesses can remain open for as long as possible. Second, this ensures that the forecasts at the time of the evacuation decision are as accurate as possible. Therefore, our models seek to recommend an optimal evacuation plan as well as the latest possible time at which it is safe to enact this plan.
